---
title: "Time_Series"
author: "Joseph Rickert"
date: "March 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BASIC TIME SERIES FORECASTING WITH R
This module provides an introduction to time series forecasting using some of the high level functions frovided by R.

### Load the packages required
```{r}
library(dplyr)             # For data manipulation
library(reshape2)          # Also for data manipulation
library(xts)               # For creating time series objects
library(forecast)          # For automatic forecasting
library(ggplot2)           # For visualization
```  
### Fetch the data and shape it for analysis
We will use the IBM stock data that we downloaded from Hahoo Finance and saved in a .csv file in a previous module.   
```{r}    
# Read the IBM data from the .csv file
dir <- "C:/DATA/IBM"
fileName <- "IBM.stock.csv"
file <- file.path(dir,fileName)

IBM <- read.csv(file)
head(IBM,2)
tail(IBM,2)

# Filter out data before 2001
IBM2 <- filter(IBM, as.Date(Date) > as.Date('2000-01-01'))
# Select only Date and Closing Price
IBM2 <- select(IBM2,Date,Close)
head(IBM2,2)
sapply(IBM2,class)
# Make Date into a Date object
IBM2 <- mutate(IBM2, Date = as.Date(Date))
# Sort in ascending order
IBM2 <- arrange(IBM2,Date)
head(IBM2,2)
```
### Create a Time Series Object
We formulate the IBM data as a time series object and then extract a small set of the data to use for training a forecasting model. We will not use the entire series to train the model so we can compare the forecasted prices with the actual recorded values.
```{r}
# Build a time series object
IBM_ts <- xts(IBM2$Close,order.by=IBM2$Date)
head(IBM_ts)
plot(IBM_ts)

# Narrow down the time series for analysis
IBM_ts2 <- IBM_ts['2012-01/2013']
plot(IBM_ts2)
```  
### Fit an ARIMA Model and Forecast
We use the functions int the forecast package to futomatically fit an Autoregressive Integrated Moving Average model and generate a forecast.   
```{r}
# Automatically fit an ARIMA Model
fit <- auto.arima(IBM_ts2)
summary(fit)

# Generate and automatic forecast
fore <- forecast(fit)
plot(fore)
fore
```
### Plot the Data with the Forecast
We build a data frame to containg the forecasted results along with the relevant portion of the actual IBM stock price data.
```{r}
# Build a common data frame for forecast
Jan2014<- IBM_ts['2014-01']
Jan_data <- as.data.frame(Jan2014)
row.names(Jan_data)
res <- data.frame(row.names(Jan_data)[1:10],Jan2014[1:10,1],
                  fore$mean,fore$lower[,1],
                  fore$upper[,1],
                  row.names = NULL)
names(res) <- c("Date", "Actual","Point","Lower","Upper")
res


meltdf <- melt(res,id="Date")
p <- ggplot(meltdf,aes(x=Date,y=value,colour=variable,group=variable)) 
p +  geom_line() + geom_point() +
  ylab("Closing Prices") +
  ggtitle("Forecast for IBM Stock")
```   
We have just scratched the surface of R's time series capabilities. For an over of all that is available in R look at the CRAN Task View: Time Series Analysis which can be found here: https://cran.r-project.org/web/views/TimeSeries.html

For a complete introductory time series course usng R seee the online book from  Rob J Hyndman and George Athanasopoulos: https://www.otexts.org/fpp
